### 2014-04-16 doc/uist14/interaction.tex
pushed my changes of figures
added comments for intro, interaction, iteration1/2

Our augmentation to Google Glass adds a 940nm 5mm IR emitter\footnote{In the prototype, we have used the IR emitter SFH 4545 from OSRAM Opto Semiconductors Inc.} (see Figure~\ref{fig:glass}). \ben{add description about the microcontroller.} In the following description, we refer to Glass because the term is succinct; but other suitable head-worn computing devices could be substituted.

@@

Though IR signal has enabled us achieve head orientation-based targeting (simplex communication), we envision that more complicated applications could be built on top. Therefore, we added 802.15.4 radio to Glass and all the target nodes.

\begin{figure}[t]
\centering
\includegraphics[width=0.9\columnwidth]{figures/GlassWithHolder.jpg}
\caption{Google Glass augmented with a repositionable IR holder.}
\label{fig:glass}
\end{figure}

@@

{\bf Scan:} The user first scans the environment to locate the position of the target. During this stage, Glass constantly sends out IR signals, and  targets offer immediate feedback when they receive a signal. The user confirms his desire to connect to a target by tapping on the Glass touchpad. Glass collects the responses from targets that have received IR reception (we use an XBee-based backchannel for this traffic). If there is only one single target in IR range, it is automatically selected. However, in a dense environment where multiple targets are within range, the user needs to perform a refinement.

@@

The overall target acquisition time thus depends on scan and refine times; the probability that refinement is needed; and the time to commit an action (tap):
\begin{equation}
t_{total}=t_{scan}+P(refine)*t_{refine}+t_{commit}
\end{equation}

>>>>

Our augmentation to Google Glass adds a 940nm 5mm IR emitter (SFH 4545 from OSRAM Opto Semiconductors Inc.). The emitter is controlled by an additional microcontroller which communicates with Google Glass through Bluetooth radio. (see Figure~\ref{fig:glass}). In the following description, we use the succint term ``Glass'' to refer to this augmented hardware.

@@

While we use IR signal for targeting, wireless communications are needed to efficiently transmit the reception status of IR signal from targets to Glass. We have chosen a commercial off-the-shelf ZigBee implementation (XBee based on 802.15.4 radio) for this purpose.

\ben{no mention of target hardware yet. I have checked in an image of the targets ``targets.JPG'' in case we need.}

\begin{figure}[t]
\centering
\includegraphics[width=1\columnwidth]{figures/GlassWithArduino.jpg}
\caption{Our Glass hardware: Google Glass augmented with a repositionable IR holder, and an additional microcontroller that communicates with Google Glass and controls IR emitter.}
\label{fig:glass}
\end{figure}

@@

{\bf Scan:} The user first scans the environment to locate the position of the target. During this stage, Glass constantly sends out IR signals, and  targets offer immediate feedback when they receive a signal. The user confirms his desire to connect to a target by tapping on the Glass touchpad. Glass collects the responses from targets that have received IR reception through the backchannel of XBee. If there is only one single target in IR range, it is automatically selected. However, in a dense environment where multiple targets are within range, the user needs to perform a refinement.

@@

The overall target acquisition time thus depends on scan and refine times; the probability that refinement is needed; and the time to commit an action (tap):
\begin{equation}
t_{total}=t_{scan}+P(refine)*t_{refine}+t_{commit}
\label{eq:time}
\end{equation}

### 2014-04-16 doc/uist14/intro.tex
Filled some of the contents; updated a few figures; just added content to disucssion and conclusion, but the sentences are broken and may not make any sense
added comments for intro, interaction, iteration1/2

%!TEX root = uist14.tex
\section{Introduction}
\bjoern{something happened to the figure caption font - it doesn't match the standard template. Please fix.}
%% from the swarm vision to the necessity of selection
The number of smart objects in our environment with embedded computation and communication has grown rapidly. These objects are all potential targets for interaction.
To initiate such interaction requires {\em selection} information about the object at the user's {\em locus of attention}~\cite{raskin}. \sean{this sentence is unclear to me.}

@@

In this paper, we describe the iterative development and evaluation of \systemnamenospace, an area-selection technique that can be readily implemented with small hardware changes to emerging head-worn devices. We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} to enable infrared (IR) communication between Glass and target appliances. In our {\em Naive IR} technique, the cone of light emitted by an IR LED (a diameter of 60-120cm and distance up to 6m) provides the {\em coarse} selection area. To {\em refine} selection when multiple targets have received IR signals, we introduce a manual list selection disambiguation technique. A study with $14$ participants compares acquisition times for physical targets in a room for our technique and an alternative list selection interface without any IR targeting. We find that target acquisition through head orientation is preferred by users and is faster than list selection, but refinements -- if necessary -- are slow.

In response, we designed an improved refinement technique, {\em Intensity IR}, in which target objects compare received IR signal strength. This value allows the system to either automatically select the target with the strongest intensity, or to re-order the refinement interface's list by their intensity values. A second study with $10$ participants shows that {\em Intensity IR} disambiguation successfully reduces both the chance of needing to do refinement as well as the time spent in list navigation when compared to {\em Naive IR}.

A final design addresses the lack of a natural mapping when users select a target in the refinement step using their device's touchpad - the axes of motion do not map directly to the spatial layout of target devices in a room. Our {\em Head-motion Refinement} technique first infers a spatial data structure of target locations using Glass' orientation sensors. In a refinement step, users can then perform head movements to change selections to spatially adjacent targets. For example, nodding down to select the target below current selection, or tilting right to select next target on the right. We present initial feedback on this technique.

@@

>>>>

%!TEX root = uist14.tex
\section{Introduction}
% \bjoern{something happened to the figure caption font - it doesn't match the standard template. Please fix.}
% \ben{fixed. Problem is because of using caption package}.

%% from the swarm vision to the necessity of selection
The number of smart objects in our environment with embedded computation and communication has grown rapidly. These objects are all potential targets for interaction. To initiate such interaction, a user needs to first acquire the target object -- a fundamental task in an interaction process that has been extensively studied in the digital world, but yet not well-explored in the physical spaces.

%% To initiate such interaction requires {\em selection} information about the object at the user's {\em locus of attention}~\cite{raskin}. \sean{this sentence is unclear to me.}

@@

In this paper, we describe the iterative development and evaluation of \systemnamenospace, an area-selection technique that can be readily implemented with small hardware changes to emerging head-worn devices. We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} to enable infrared (IR) communication between Glass and target appliances. In our {\em Naive IR} technique, the cone of light emitted by an IR LED (a diameter of 30-60cm and distance up to 8m) provides the {\em coarse} selection area (illustrated in Figure~\ref{fig:teaser}A, where targets that have received IR signal light up). To {\em refine} selection when multiple targets have received IR signals, our solution starts with a naive ordering -- based on their names. This is shown in Figure~\ref{fig:teaser}B, where the targets in range are ordered numerically on the near-eye display. A study with $14$ participants compares acquisition times for physical targets in a room for our technique and an alternative list selection interface without any IR targeting \ben{the ordered-list is also list selection, needs clarification!} . We find that target acquisition through head orientation is preferred by users and is faster than list selection but refinements -- if necessary -- are slow.

In response, we designed an improved refinement technique, {\em Intensity IR}, in which target objects compare received IR signal strength. This value allows the system to either eliminate some distracting targets on the peripheral, or to re-order the refinement interface's list by their intensity values. For example, in Figure~\ref{fig:teaser}B of {\em Intensity IR} technique, 5 is eliminated first and the list is re-ordered based on the intensity readings. A second study with $10$ participants proves that {\em Intensity IR} successfully reduces both the chance of needing to do refinement as well as the time spent in list navigation when compared to {\em Naive IR}.

A final design addresses the lack of a natural mapping when users select a target in the refinement step using their device's touchpad - the axes of motion do not map directly to the spatial layout of target devices in a room. Our {\em Head-motion Refinement} technique first learns the relative spatial structure of the targets using Glass' orientation sensors. Then in future target acquisition stage, users can then perform head movements to change selections to spatially adjacent targets (Figure~\ref{fig:teaser}C). For example, nodding down to select the target below current selection, or tilting right to select next target on the right. We present initial feedback on this technique.

@@

\ben{I am considering adding back the contribution part, because the paper itself mainly focuses on the iterative design process. Without an explicit contribution statement, the audience will have to infer and summarize themselves.}

### 2014-04-16 doc/uist14/iteration1.tex
added comments for intro, interaction, iteration1/2

\section{Iteration 1: Naive IR}
\subsection{Problem Statement and Goal}
Our initial research question is: {\em Can IR-based targeting reduce the selection time compared to UI list navigation on a head-worn device?}

\subsection{Technique}
In our first implementation, we use IR for scanning as described in the previous section. For the refinement technique, we simply show a list of the subset of targets that have received IR signals on the Glass display and let users swipe to select from that list and tap to confirm.

A natural point of comparison is an interface that does not use any orientation information - it always shows 
a complete list of all targets. We implemented a list view where users swiped forwards and backwards on the Glass touchpad to navigate incrementally through the list. To quantify the benefit of using IR for the scanning step, we carried out a target acquisition study which compares the {\em Naive IR selection} and {\em list selection}.

@@

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/targeting-study-layout.pdf}
\caption{In the targeting study, participants were asked to find and select one of 10 targets in the lab environment. \ben{add a way to describe the distance in the figure}.}
\label{fig:targeting-study-layout}
\end{figure}

In the within-subject study, half of the participants performed {\em IR selection} first and the other half used {\em list selection} first. For each selection condition, we conducted 15 target selection by randomly choosing from all the targets. During the study, we measure the {\bf target acquisition time} for each target selection. Afterwards, participants were asked to complete a survey that elicited answers to Likert-scale questions as well as open-ended answers about their experience.

@@

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/result_study1a.pdf}
\caption{\ben{fix this later}. Boxplot of task completion times for the comparison between {\em infrared} mode and {\em list} mode (A), and between IR multiple responses cases and IR single response cases (B). The centers of boxes are median values, while white dashed lines are mean values.}
\label{fig:ir_vs_list}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=1.09\columnwidth]{figures/result_study1b.pdf}
\caption{With and without refinement.}
\label{fig:with_without_refinement}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/result_study1c.pdf}
\caption{Times taken to select a device vs.~its order in the list. The dotted line is a linear fit between the mean times and device orders. Two horizontal lines of mean target acquisition times in {\em infrared} mode are also annotated for comparison.}
\label{fig:time-vs-list-order}
\end{figure}

Our results indicate that {\em IR selection} outperforms {\em list selection}.  The average target acquisition time for {\em IR selection} is 6.67 seconds, while {\em list selection} took 8.86 seconds (see Figure~\ref{fig:ir_vs_list}). Student's t-test shows a significant difference ($t(279)=-3.81, p=0.00017$).

@@

In summary, {\em Naive IR} can outperform linear list selection, but the refinement step detracts significantly from the efficiency of the technique.
% learned that with IR, we can have effectively reduced the overall target acquisition time. And the major gain comes from the fact that {\em coarse selection} helps reduce the chances of entering {\em fine selection} stage.

\ben{qualitative feedback hasn't been added yet for this part.}
%\subsubsection{Preference}

>>>>

\section{Iteration 1: Naive IR}
\subsection{Problem Statement and Goal}
Our initial research question is: {\em Can IR-based targeting reduce the selection time compared to the case where only UI list navigation is used on a head-worn device?}

\subsection{Technique}
In our first implementation, we use IR for scanning as described in the previous section. For the refinement technique, we simply show a list of the subset of targets that have received IR signals on the Glass display. Since we only know that the targets are within the range of users' head orientation. We order them using their ids; and in realistic environment target names can also be used if names can be assigned. We then let users swipe to select from that list and tap to confirm the intended target.

A natural point of comparison is an interface that does not use any head orientation information - it always shows a complete list of all targets. We implemented a list view where users swiped forwards and backwards on the Glass touchpad to navigate incrementally through the list. To quantify the benefit of using IR for the scanning step, we carried out a target acquisition study which compares the {\em Naive IR selection} and {\em list selection}.

@@

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/study-layout1.png}
\caption{In the targeting study, participants were asked to find and select one of 10 targets in the lab environment.}
\label{fig:targeting-study-layout}
\end{figure}

In the within-subject study, half of the participants performed {\em IR selection} first and the other half used {\em list selection} first. For each selection condition, we conducted 15 target selection by randomly choosing from all the targets. During the study, we measured the {\bf target acquisition time} for each target selection. Afterwards, participants were asked to complete a survey that is mainly open-ended answers about their experience.

@@

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/result_study1a.pdf}
\caption{Boxplot of target acquisition time for {\em IR selection} and {\em list selectioN} is shown on the left. The center of each box is the median value, while we annotate the mean value using white dashed lines.}
\label{fig:ir_vs_list}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=1.09\columnwidth]{figures/result_study1b.pdf}
\caption{Boxplot and CDF of target acquisition time when refinement is needed or not in {\em IR selection}}.
\label{fig:with_without_refinement}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/result_study1c.pdf}
\caption{Times taken to select a device vs.~its order in the {\em list selection}. The dotted line is a linear fit between the average time and target orders in the list. Two horizontal solid lines are the average target acquisition times in {\em Naive IR} when refinement is needed or not.}
\label{fig:time-vs-list-order}
\end{figure}

Our results indicate that {\em IR selection} outperforms {\em list selection}.  The average target acquisition time for {\em IR selection} is 6.67 seconds \ben{one concern that I have is this is extremely long, 6 seconds. Do we need to explain more? One possible reason here is that we ask the user not to look at targets until we called out, so the visual search time is around 3 seconds. Should we say more or leaving 6 seconds there is fine?}, while {\em list selection} took 8.86 {\em list selection also requires visual search, I think...} seconds (see Figure~\ref{fig:ir_vs_list}). Student's t-test shows a significant difference ($t(279)=-3.81, p=0.00017$).

@@

As for the qualitative feedback, 11 out of 14 participants preferred IR selection over list mode (two preferred list, one was undecided). While both interfaces were judged similarly on overall ease of connecting, list selection was perceived to be cumbersome. One user noted benefit of {\em IR selection} being \studyquote{more direct [than list mode]}, allowing users to focus on the targeted objects instead of the screen. One subject called it \studyquote{natural to interact with things just by looking at them}. Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}, and many participants compared the overall time of target acquistion. One subject pointed out that \studyquote{{\em IR selection} doesn't requires you to swipe a lot of times like list mode but list selection}.

In summary, {\em Naive IR} can outperform linear list selection, and most users like this head-orientation based targeting. But from our quantified study result, the refinement step detracts significantly from the efficiency of the technique.

% learned that with IR, we can have effectively reduced the overall target acquisition time. And the major gain comes from the fact that {\em coarse selection} helps reduce the chances of entering {\em fine selection} stage.

%\subsubsection{Preference}

### 2014-04-16 doc/uist14/iteration2.tex
added comments for intro, interaction, iteration1/2

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/study-layout2.jpg}
\caption{The current environment setup \ben{needs annotation and a better photo.} \bjoern{yes}.. blackboard is 2*107+122 cm , 183 cm dimension.}
\label{fig:study-layout2}
\end{figure}

@@

%% resolved: \bjoern{for consistency's sake, can we show a box plot as in figure 5?} \ben{I think cdf is a better way than boxplot? It actually provides more information. I may consider using cdf consistently.}

@@

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/result_study2.pdf}
\caption{The target acquisition of in {\em Naive IR} and {\em Intensity IR} condition.}
\label{fig:study2}
\end{figure}

One side effect that we have observed in this approach is that the {\em Intensity IR} sometimes eliminates the desired target during the {\em scanning} stage. The this error rate (13 out of 300 cases, 4.3\%), is higher than the error rate for {\em Naive IR} \bjoern{insert error rate}. % but this is the trade-off for the overall faster acquisition time.

>>>>

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/study-layout2.png}
\caption{The environment setup for our second study. In comparison to our first study, we have delibrately increased the target density.}
\label{fig:study-layout2}
\end{figure}

@@

@@

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/result_study2.pdf}
\caption{Boxplot and CDF of the target acquisition time in {\em Naive IR} and {\em Intensity IR} conditions.}
\label{fig:study2}
\end{figure}

One side effect that we have observed in this approach is that the {\em Intensity IR} sometimes eliminates the desired target during the {\em scanning} stage. Out of 300 trials, 13 of them (4.3\%) has been accidentally eliminated. This is higher than the error rate for {\em Naive IR} (only 5 out of 300, ~1.6\%). % but this is the trade-off for the overall faster acquisition time.
